### 2016-01-20 numbat_pres/numbatpres.Rmd
slides for NUMBAT

---
title: "Bushfires in R"
author: "Elle Saber"
date: "20 January 2016"
output: beamer_presentation
---

## Modelling Probability of Ignition of Bushfires in Victoria

Data Processing involved with the honours project

## Slide with Bullets
```{r mysize=TRUE, size='small'}
library(ncdf)
temp.nc <- open.ncdf("./TMAX1984-2009.nc")
temp.nc

```

## Fire Map

![Map of Fire Occurence](figures/Data/firemap.pdf)

## NCDF Temp Data random Day

![Max Temperature](figures/Data/tmax.pdf)

## NCDF time series

![Time series Silvan](figures/Data/silvan.pdf)